---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# tukeygrps

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/tukeygrps)](https://cran.r-project.org/package=tukeygrps)
<!-- badges: end -->

Tukeygrps provides a simple helper function for the letter annotation of (gg)plots according to statistical differences between groups determined by a Tukey-HSD test.

## Installation

You can install tukeygrps from github using devtools:

``` {r, eval=FALSE}
install.packages("devtools")
install_github("leonardblaschek/tukeygrps")
```

## Example

Here we use tukey_groups() to quickly add letters to a ggplot. Alpha is set to 0.001, and the resulting letters indicate statistical differences within each panel.

```{r example, warning=FALSE, message=FALSE}
library(tukeygrps)
library(tidyverse)

data(mpg)
head(mpg)

tukey_letters <- tukey_groups(mpg, hwy, class, 0, 0.001, cyl)

head(tukey_letters)

ggplot() +
  geom_point(data = mpg, 
             aes(x = class,
                 y = hwy)) +
  geom_text(data = tukey_letters,
            aes(x = class,
                y = hwy,
                label = groups)) +
  facet_wrap(~ cyl) +
  coord_flip()

```

**Note:** The distributions and sample numbers in this example dataset should discourage anyone from applying a parametric test such as the Tukey-HSD. Before you use this function, make sure that your observations are:

* normally distributed
* homoscedastic
* independent within and between groups
* of similar sample size
